---
title: "Trail Difficulties at MacLeish Field Station"
author: "Lynn Albright and Maggie Szlosek"
date: "15 April 2018"
output:
  html_document:
    code_folding: hide
    theme: cosmo
  pdf_document: default
  
---

```{r include=FALSE, cache=FALSE}
library(tidyverse)
library(macleish)
library(leaflet)
library(sf)
```

```{r message=FALSE,warning=FALSE}
# INITIALIZING DATA
contours3m <- macleish_layers [["boundary"]] %>%
  st_intersection(macleish_layers[["contours_3m"]])
trails <- macleish_layers[["trails"]]
porc <- trails %>% filter(name=="Porcupine Trail")
west <- trails %>% filter(name=="Western Loop")
vern <- trails %>% filter(name=="Vernal Pool Loop")
east <- trails %>% filter(name=="Eastern Loop")
eout <- trails %>% filter(name=="Easy Out")
road <- trails %>% filter(name=="Poplar Hill Road")

# SETTING THE CRS EQUAL
thecrs <- st_crs(macleish_layers[["trails"]], asText = TRUE)
bound <- st_union(macleish_layers[["boundary"]])
macmap <- st_convex_hull(bound)

# CONTOURS AND ROAD
macmap <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = contours3m, weight = 1,color="darkblue") %>%
  addPolylines(data = road, weight = 5, color="grey",opacity=0.75) %>%
  addPolygons(data = bound, weight = 2, color="green",opacity=1, fillOpacity = 0.1) %>%
  
# TRAILS
  addPolylines(data = porc, weight = 3, color="magenta",opacity=0.75) %>%
  addPolylines(data = west, weight = 3, color="orange",opacity=0.75) %>%
  addPolylines(data = vern, weight = 3, color="yellow",opacity=0.75) %>%
  addPolylines(data = east, weight = 3, color="red",opacity=0.75) %>%
  addPolylines(data = eout, weight = 3, color="purple",opacity=0.75) %>%

# LEGEND
  addLegend("bottomleft", 
    colors =c("orange",  "red", "yellow", "magenta", "purple","darkblue","grey"),
    labels= c("West Loop", "East Loop","Vernal Pool Loop","Porcupine Trail", "Easy Outs for West Loop","Contour Lines (Not a Trail)","Poplar Hill Road"),
    title= "MacLeish Field Station Trail Guide",
    opacity = 1)

```
```{r message=FALSE,warning=FALSE}
# CALCULATING INTERSECTIONS
#The "sum" function arguments taken partially from 
#stackoverflow.com/questions/2190756/how-to-count-true-values-in-a-logical-vector
#Many thanks to Stack Overflow user "Marek".
vernint <- sum(st_intersects(contours3m,vern,sparse=FALSE), na.rm=TRUE) # 
westint <- sum(st_intersects(contours3m,west,sparse=FALSE), na.rm=TRUE)
eastint <- sum(st_intersects(contours3m,east,sparse=FALSE), na.rm=TRUE)
porcint <- sum(st_intersects(contours3m,porc,sparse=FALSE), na.rm=TRUE)

#CALCULATING TRAIL LENGTHS
westlen <- sum(st_length(west))
eastlen <- sum(st_length(east))
vernlen <- sum(st_length(vern))
porclen <- sum(st_length(porc))

# CALCULATING DIFFICULTY LEVELS
westfin <- sum(westint*4, westlen)
vernfin <- sum(vernint*4, vernlen)
eastfin <- sum(eastint*4, eastlen)
porcfin <- sum(porcint*4, porclen)

# DRAW FINAL MAP
macmap
```                                  
<br>
Macleish Field Station has a variety of resources available to members of the Smith community, among them five trails for visitors to wander. However, MacLeish currently has no system in place to rate the difficulty of a trail, which means that visitors have no way of knowing how long walking the trail will take, if they are physically capable of completing the full trail, or if they would rather take one of the other trails that is more suited to their personal needs. In order to rectify this, we have created a formula in order to compare trails, and ranked each trail based on the numerical difficulty value it has been given. The full project can be found on Github.[^1]  

[^1]: [Here is the full project on Github](https://github.com/lwalbright/miniproj3)


The ranking is based on two factors, the length of the trail and the elevation change of the trail. Elevation was computed from the MacLeish contour map data, in which each line indicates a 3 meter change in elevation. We calculated "elevation" as the total number of contour lines crossed by each trail. Because a steep walk is generally harder than a long walk, we decided to weight the elevation significantly against the length. Lengths were calculated from the MacLeish trail map data. The difficulties were calculated as follows:         

$difficulty = length + 4 * elevation$   
                
From this formula, we get the following difficulty scores: 


Trail Name    | Difficulty Score
------------- | -------------
West Loop     | `r round(westfin, 2)`
East Loop     | `r round(eastfin, 2)`
Vernal Pool Loop | `r round(vernfin, 2)` 
Porcupine Trail | `r round(porcfin, 2)` 
                 
We then took these scores and assigned difficulties to them. "Difficult"" is any score greater than 1500, "Moderate" is any score greater than or equal to 500, and "Easy" is any score less than 500. By applying this formula, we found the following:          
              
Trail Name    | Difficulty
------------- | -------------
West Loop     | Moderate
East Loop     | Difficult
Vernal Pool Loop | Easy
Porcupine Trail | Moderate

                    
The formula, however, is not perfect. Particularly with the East Loop, the trail follows and repeatedly crosses the same contour line, which indicates a relatively flat section of the trail. Our calculations did not control for repeated crossings of the same line (as that would, in most cases, be important to count). This did have the result of skewing our data somewhat. However, the East Loop Trail is also the longest trail, and includes several steeper sections, which makes it our hardest trail regardless. Furthermore, East Loop is the only trail to feature "Easy Outs", which allow hikers to shorten their hike by cutting out large sections of the trail. If visitors choose to make use of these features, they would find the East Loop considerably easier.            

Providing hikers with this information shouldn't be particularly hard, as the only thing that would need to be added are posted signs with the difficulty levels and a brief overview of each trail. The cost and time required to add these signs would be negligible compared to their potential benefit. These signs could also include any noteworthy points of interest on each trail, the average slope, and many other details that were likely calculated by other project groups. Assuming MacLeish already has posted signs, an additional sign could easily be appended to each signpost giving further trail information. The cost would depend largely on the type of weatherproofing MacLeish decides to apply to these signs. The goal should be to maximize the amount of easily understandable information available to MacLeish visitors.                

This solution would not take very long to implement. Even including the time it would take to design the new signs, they could be made and posted within a fairly short timeframe. MacLeish Field Station could easily install temporary laminated signs on printer paper in the meantime. The environmental impact would depend largely on how MacLeish decides to make and display the signs. Obviously they would need to consider the environmental safety of adding such a sign, as any modifications to the natural landscape can affect the flora and fauna therein. After the signs are made and their locations decided, installation would be a simple (but worthwhile) endeavor.                         

TODO: 
remove TODO section
split up code chunks
be super anal about code neatness
add writeup
  EDIT writeup
  clear concise desc of data used
  reporting at least one actionable recommendation (what should Smith DO with your findings?)
  a short discussion on the resources necessary to implement your recommendation:
footnote with Github
issues tab?
  

